\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-25 draft.tex
Changes in Section 2

For presentation purposes, we provide a sketch of a highly simplified language
for describing reducers. The language allows us to give examples of reducers and
to discuss our approach for checking their commutativity. Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, their combinations over basic
arithmetic operations, or a non-deterministically chosen integer value $*$. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly for each input element until
the entire input list is consumed. An example of a reducer is shown in Figure~\ref{fig:reducer_example}. To simplify presentation, we assume in this paper the reducer never output in the loop body $s_2$. In the J-ReCoVer tool, an algorithm to deal with output inside the loop is implemented.

Some reducers use two (or more) top-level loops to compute the output, possibly
interleaved with some non-looping code. These loops are executed sequentially,
repeatedly iterating over the input list from its beginning. For example, for
calculating the standard deviation, one first computes the average of inputs and
then uses it to compute the final result (using two passes over the input list).
In that case, we suggest to verify the reducer by first partitioning it into two
(or more) reducers, each containing a single top-level loop, and then verifying
these reducers separately.\footnote{Nested loops can be removed by adding
additional branch statements since both the inner and outer loop are over the
same input list. In fact, such a program construct has never occurred in the
examples we have seen.} The top-level loops communicate through shared
variables. After the transformation, reducers corresponding to the second
top-level loop (and possibly further such loops) will work with random initial
values of the shared variable, which over-approximates the original behaviour.
Usually, the second (or further) top-level loop are commutative even with
arbitrary initial shared variable values and hence J-ReCoVer can be used to
prove it.

>>>>

For presentation purposes, we provide a sketch of a highly simplified language
for describing reducers. The language allows us to give examples of reducers and
to discuss our approach for checking their commutativity. Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, their combinations over basic
arithmetic operations, or a non-deterministically chosen integer value $*$. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly for
each input element until the entire input list is consumed. An example of a
reducer is shown in Figure~\ref{fig:reducer_example}. In the paper, to simplify
the presentation, we assume that a reducer does never produce output in the loop
body $s_2$. In the J-ReCoVer tool, an algorithm to deal with output inside the
loop is implemented.

% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%
% Just to remember: the way the output from the main loop is handled currently
% is wrong: One should either check that there is no flow between the input and
% output in the loop, or to add a check that the projection of the F function on
% any value output in the main loop stays the same for any input value.
%
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

Some reducers use two (or more) top-level loops to compute the output, possibly
interleaved with some non-looping code. These loops are executed sequentially,
repeatedly iterating over the input list from its beginning. For example, for
calculating the standard deviation, one first computes the average of inputs and
then uses it to compute the final result (using two passes over the input list).
In that case, we suggest to verify the reducer by first partitioning it into two
(or more) reducers, each containing a single top-level loop, and then verifying
these reducers separately.\footnote{Nested loops can be removed by adding
additional branch statements since both the inner and outer loop are over the
same input list. In fact, such a program construct has never occurred in the
examples we have seen.} The top-level loops communicate through shared
variables. After the transformation, reducers corresponding to the second
top-level loop (and possibly further such loops) will work with random initial
values of the shared variable, which over-approximates the original behaviour.
In our experience, the second (or further) top-level loop are usually
commutative even with arbitrary initial shared variable values, and so J-ReCoVer
can be used to handle such reducers.
